---
title: "Mock Analysis of Fitbit User Data"
output: html_document
runtime: shiny
--- 
# Analysis of Fitbit User Data

## Ask
This project is a mock analysis performed for Bellabeat, a company that sell products for women's health. In this scenario, Bellabeat has commissioned an analysis to be performed on smart device usage data to help inform their marketing team. The team intends to use the insights gained from this analysis to guide them in developing a marketing strategy for a remote fitness tracker device.

## Prepare
CEO of Bellabeat Urška Sršen directed the analysis to be performed on the public Fitbit dataset [FitBit Fitness Track Data](https://www.kaggle.com/datasets/arashnic/fitbit) at Kaggle. The data is also housed in other online locations--this analysis uses the datasets provided through [Zenodo](https://zenodo.org/record/53894#.Y_-BtHbMKCp) because it contains additional information collected during the latter of the two months that the data collection period spans. The data shared between the Zenodo and Kaggle datasets are otherwise identical in structure and contents, as shown by checksum (provided in cs.js) and grep tests.  

The authors of the data include [Robert D. Furberg, Ph.D.,](https://www.researchgate.net/scientific-contributions/Robert-D-Furberg-32290751) and three other associates (Brinton, Keating, & Ortiz.) They generated the information in the datasets by monitoring thirty Fitbit users over a period of two months. This information includes a range of metrics that can be remotely detected or inferred through the Fitbit device, such as periods of activity, the intensity of the activity, heart rate, periods of sleep, and calorie expenditure. These measurements are organized by day, hour, and minute. The information is also supplemented by user-reported data such as height, weight, gender, and age, and some measurements, such as calorie expenditure, are derived from these metrics.

### Limitations
There are numerous issues with the data that limit its ability to generalize insights to the population at large. For one, only thirty participants were tracked for this program. This sample size barely meets the threshold that many consider to be suitable for applying the insights to the population at large. Another issue is that the sample is very likely to self-selecting--participants are likely to be younger and [able to spare the expense of owning a Fitbit](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5427248/) and other health amenities. Participants were crowdsourced through the Amazon Mechanical Turk marketplace, where they were compensated for their participation ($10). The framing of the data is also slightly complicated by the fact that the original study was to assess the effectiveness of using remote tracking devices and not necessarily in providing quality health or marketing data.

There is also a glaring problem with the data in how it specifically relates to women. Not only are women a subset of the relatively small sample (the information is collected from two genders rather than exclusively women), it doesn't make any distinctions that highlight contemporary cultural and demographic shifts, such as participants who might identify as a woman but whose sex is male, or whether or not participants are undergoing hormone therapy. Such individuals are very likely to have been excluded in the data collection process and it will be very difficult to apply with certainty any insights collected to their health needs and goals.

Additionally, a number of smaller issues impact the data's reliability. These issues include the degree of self-reporting, a lack of direct engagement with researchers, a lack of insight into the participant eligibility process, the fact that the researchers were panelled (meaning the researchers chose the participants, ostensibly for variation), the possible use of different Fitbit devices (each of which may have different metrics capabilities and sensor fidelities), and that users may manually override at least some data reported. The categories of fitness intensities are also possibly arbitrary--as originally decided by Fitbit, with distinctions hidden behind the manufacturer's proprietary algorithms. Finally, there are datasets reported to exist in [Fitabase](https://www.fitabase.com/media/1930/fitabasedatadictionary102320.pdf) (a third-party device data usage collector) that does not exist in the published datasets, including a food log.

Finally, there is a problem with the period that the data spans, which is only two months. Fitness, and especially weight-loss, are long-term goals, with a person's weight varying naturally week-to-week. A comprehensive analysis of participants being able to meet their weight-loss goals would require a longer study period.

Still, insights from the data could be worthy of consideration. The volume of information is quite substantial and at least minimally represents *some* groups in the general population.

## Process
This analysis uses the Tidyverse suite for the R programming language to clean and filter the data. This section is technical and pedantic--the average reader may want to skip to the results of analysis.

### Validation
The first step is to validate all of the fields in all of the datasets. The fields should not contain extreme, incomprehensible, or improper values, with consistent formatting for strings.

#### date and datetime, N/A fields
The **date** fields should have a format of mm/dd/yyyy and the **datetime** fields should contain a format of 'mm/dd/yyyy hh:mm:ss'. All of the datasets have a label for a time record. The transformation from string to datetime happens as the datasets are loaded. An invalid interpretation of the string would produce a voided (N/A) field. A cursory sweep of the loaded-in datasets shows that the only voided fields in any column are found in the weight logs (aside: since the weight logs are user-reported, this may also attest to the fidelity of the device.) Almost none of the participants in this study measured their body fat percentage, and the only one who did so more than once (subject 1503960366) observed no change.

#### additional data, duplication
The second month of the study has additional datasets, which include:
1. minuteCaloriesWide_merged.csv,
1. minuteIntensitiesWide_merged.csv.csv,
1. and minuteStepsWide_merged.csv,

These were assumed to be pre-transformation versions of existing narrowed versions in the final draft; However, once narrowed, each is shown to contain a different number of entries from their original counterpart.

* m2_minute_calories_n has 04-12, missing 05-13, plus other unknowns
* m2_minute_calories_n_v2 missing 04-12, has 05-13, plus other unknowns
* minute_intensities_* and minute_steps_* likewise share the same issue

This suggests the original and additional data may each contain information missing from the other. Tests show the differences are not the result of duplication, so the datasets have been unioned.

### Outlier values
Summaries show several very strong performances, but they appear to be confined to the limits of human capability (for example, burning 4528 calories per day rather than 45,280.) This is likely a result of the fidelity of automated remote tracking.

### Final Compilations
The original datasets for both months have a *Daily Activity* dataset, which is a collage of data taken from *Calories*, *Steps*, and *Intensities*. Both of these datasets have been rebuilt for this analysis, incorporating the additional data found on Zenodo (note: these collages are built **only** from datasets that provide minute-level measurements.) The two collages are then unioned. The final collage includes information on METs, but excludes information on distance traveled. Despite being defined in the [Fitabase dictionary](https://www.fitabase.com/media/1930/fitabasedatadictionary102320.pdf), this information is not available in other parts of the published datasets except for in the original daily activity collages. However, the number of steps and workout intensities may be good substitutes, and may also be less likely to have misleading data (for example, that generated from a high-intensity run on a treadmill.)

**(Note on the rendering: The plots generated in this analysis were originally tuned for rendering in Visual Studio Code and may appear different on other platforms.)**

```{r}
library(tidyverse)
library(stringr)
options(scipen = 999)
options(warn=-1)

# Month one and month two share:
#  - dailyActivity_merged.csv
#  - heartrate_seconds_merged.csv
#  - hourlyCalories_merged.csv
#  - hourlyIntensities_merged.csv
#  - hourlySteps_merged.csv
#  - minuteCaloriesNarrow_merged.csv
#  - minuteIntensitiesNarrow_merged.csv
#  - minuteMETsNarrow_merged.csv
#  - minuteSleep_merged.csv
#  - minuteStepsNarrow_merged.csv
#  - weightLogInfo_merged.csv

# Shared files must be checked for consistency
compare_tibbles <- function(t1, t2, set_index) {
    t1_col_names <- colnames(t1)
    t2_col_names <- colnames(t2)

    pass_flag <- TRUE
    if (length(t1_col_names) != length(t2_col_names)) {
        print(str_glue("Tibbles in set {set_index} do not have the same number of columns.")) 
        pass_flag <- FALSE
    }

    for (i in seq_along(t1_col_names)) {
        if (is.na(t2_col_names[i])) {
            print(str_glue("Tibbles 2 in set {set_index} is missing column {t1_col_names[i]}")) 
        } else if (t1_col_names[i] != t2_col_names[i]) {
            print(str_glue("Column names in tibbles pair {set_index} are not identical: {t1_col_names[i]} | {t2_col_names[i]}")) 
            pass_flag <- FALSE
        }
    }

    # Actually, this is more fancy and direct...
    str_t1 <- capture.output(str(t1))
    str_t2 <- capture.output(str(t2))
    start_index1 <- grep("\\.\\. cols\\(", str_t1)
    end_index1 <- grep("\\.\\. \\)", str_t1)
    start_index2 <- grep("\\.\\. cols\\(", str_t2)
    end_index2 <- grep("\\.\\. \\)", str_t2)

    if (!identical(str_t1[start_index1:end_index1], str_t2[start_index2:end_index2])) { 
        print(str_glue("Tibbles in set {set_index} are not structurally identical!")) 
        pass_flag <- FALSE
    }

    if (pass_flag) {
        return(TRUE)
    } else {
        return(FALSE)
    }
}

#### SHARED DATA
## note: _n is used to mark data that is narrow, _w for wide

# Activity
m1_daily_activity <- read_csv(
    "datasets/fitbit_mturk_3.12.16-4.11.16/dailyActivity_merged.csv",
    col_types = cols(
        ActivityDate = col_date(format = "%m/%d/%Y")
    ),
    show_col_types = FALSE
) 
m2_daily_activity <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/dailyActivity_merged.csv",
    col_types = cols(
        ActivityDate = col_date(format = "%m/%d/%Y")
    ),
    show_col_types = FALSE
) 
daily_activity_tibbles <- list(m1_daily_activity, m2_daily_activity)

# Heart rate
m1_heartrate_seconds <- read_csv(
    "datasets/fitbit_mturk_3.12.16-4.11.16/heartrate_seconds_merged.csv",
    col_types = cols(
        Time = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
m2_heartrate_seconds <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/heartrate_seconds_merged.csv",
    col_types = cols(
        Time = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
heartrate_seconds_tibbles <- list(m1_heartrate_seconds, m2_heartrate_seconds)

# Calories (in kilocalories)
m1_hourly_calories <- read_csv(
    "datasets/fitbit_mturk_3.12.16-4.11.16/hourlyCalories_merged.csv",
    col_types = cols(
        ActivityHour = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
m2_hourly_calories <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/hourlyCalories_merged.csv",
    col_types = cols(
        ActivityHour = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
hourly_calories_tibbles <- list(m1_hourly_calories, m2_hourly_calories)

m1_minute_calories_n <- read_csv(
    "datasets/fitbit_mturk_3.12.16-4.11.16/minuteCaloriesNarrow_merged.csv",
    col_types = cols(
        ActivityMinute = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
m2_minute_calories_n <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/minuteCaloriesNarrow_merged.csv",
    col_types = cols(
        ActivityMinute = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
minute_calories_n_tibbles <- list(m1_minute_calories_n, m2_minute_calories_n)

# Intensities
m1_hourly_intensities <- read_csv(
    "datasets/fitbit_mturk_3.12.16-4.11.16/hourlyIntensities_merged.csv",
    col_types = cols(
        ActivityHour = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
m2_hourly_intensities <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/hourlyIntensities_merged.csv",
    col_types = cols(
        ActivityHour = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
hourly_intensities_tibbles <- list(m1_hourly_intensities, m2_hourly_intensities)

m1_minute_intensities_n <- read_csv(
    "datasets/fitbit_mturk_3.12.16-4.11.16/minuteIntensitiesNarrow_merged.csv",
    col_types = cols(
        ActivityMinute = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
m2_minute_intensities_n <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/minuteIntensitiesNarrow_merged.csv",
    col_types = cols(
        ActivityMinute = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
minute_intensities_n_tibbles <- list(m1_minute_intensities_n, m2_minute_intensities_n) 

# Steps
m1_hourly_steps <- read_csv(
    "datasets/fitbit_mturk_3.12.16-4.11.16/hourlySteps_merged.csv",
    col_types = cols(
        ActivityHour = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
m2_hourly_steps <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/hourlySteps_merged.csv",
    col_types = cols(
        ActivityHour = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
hourly_steps_tibbles <- list(m1_hourly_steps, m2_hourly_steps)

m1_minute_steps_n <- read_csv(
    "datasets/fitbit_mturk_3.12.16-4.11.16/minuteStepsNarrow_merged.csv",
    col_types = cols(
        ActivityMinute = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
m2_minute_steps_n <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/minuteStepsNarrow_merged.csv",
    col_types = cols(
        ActivityMinute = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
minute_steps_tibbles <- list(m1_minute_steps_n, m2_minute_steps_n)

# METs (divide these values by 10)
m1_minute_mets_n <- read_csv(
    "datasets/fitbit_mturk_3.12.16-4.11.16/minuteMETsNarrow_merged.csv",
    col_types = cols(
        ActivityMinute = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
m2_minute_mets_n <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/minuteMETsNarrow_merged.csv",
    col_types = cols(
        ActivityMinute = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
mets_n_tibbles <- list(m1_minute_mets_n, m2_minute_mets_n)

# Sleep
m1_minute_sleep <- read_csv(
    "datasets/fitbit_mturk_3.12.16-4.11.16/minuteSleep_merged.csv",
    col_types = cols(
        date = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
m2_minute_sleep <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/minuteSleep_merged.csv",
    col_types = cols(
        date = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
minute_sleep_tibbles <- list(m1_minute_sleep, m2_minute_sleep)

# Weight - User reported
m1_weight_log <- read_csv(
    "datasets/fitbit_mturk_3.12.16-4.11.16/weightLogInfo_merged.csv",
    col_types = cols(
        Date = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
m2_weight_log <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/weightLogInfo_merged.csv",
    col_types = cols(
        Date = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
weight_log_tibbles <- list(m1_weight_log, m2_weight_log)

shared_tibbles <- list(
    daily_activity_tibbles,
    heartrate_seconds_tibbles,
    hourly_calories_tibbles,
    hourly_intensities_tibbles,
    hourly_steps_tibbles,
    minute_calories_n_tibbles,
    minute_intensities_n_tibbles,
    mets_n_tibbles,
    minute_sleep_tibbles,
    minute_steps_tibbles,
    weight_log_tibbles
)

#### ADDITIONAL/SUPPLEMENTARY DATA

# Month one is missing:
#  - dailyCalories_merged.csv
#  - dailyIntensities_merged.csv
#  - dailySteps_merged.csv
#  - minuteCaloriesWide_merged.csv
#  - minuteIntensitiesWide_merged.csv
#  - minuteStepsWide_merged.csv
#  - sleepDay_merged.csv

m2_daily_calories <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/dailyCalories_merged.csv",
    col_types = cols(
        ActivityDay = col_datetime(format = "%m/%d/%Y")
    ),
    show_col_types = FALSE
)

m2_daily_intensities <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/dailyIntensities_merged.csv",
    col_types = cols(
        ActivityDay = col_datetime(format = "%m/%d/%Y")
    ),
    show_col_types = FALSE
)

m2_daily_steps <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/dailySteps_merged.csv",
    col_types = cols(
        ActivityDay = col_datetime(format = "%m/%d/%Y")
    ),
    show_col_types = FALSE
)

# Narrowed -- m2_minute_calories_n already exists, second version for comparison
m2_minute_calories_w <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/minuteCaloriesWide_merged.csv",
    col_types = cols(
        ActivityHour = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
m2_minute_calories_n_v2 <- m2_minute_calories_w %>%
    pivot_longer(
        cols = starts_with("Calories"),
        names_to = "ActivityMinute",
        values_to = "Calories"
    ) %>%
    mutate(
        ActivityMinute = ActivityHour + minutes(as.integer(substr(ActivityMinute, 9, 10))), 
        ActivityHour = NULL,
        .before = "Calories"
    )

## Narrowed -- m2_minute_intensities_n already exists, v2 for comparison
m2_minute_intensities_w <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/minuteIntensitiesWide_merged.csv",
    col_types = cols(
        ActivityHour = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
    show_col_types = FALSE
)
m2_minute_intensities_n_v2 <- m2_minute_intensities_w %>%
    pivot_longer(
        cols = starts_with("Intensity"),
        names_to = "ActivityMinute",
        values_to = "Intensity"
    ) %>%
    mutate(
        ActivityMinute = ActivityHour + minutes(as.integer(substr(ActivityMinute, 10, 11))),
        ActivityHour = NULL,
        .before = "Intensity"
    )

# Narrowed -- m2_minute_steps_n already exists, v2 for comparison
m2_minute_steps_w <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/minuteStepsWide_merged.csv",
    col_types = cols(
        ActivityHour = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
)
m2_minute_steps_n_v2 <- m2_minute_steps_w %>%
    pivot_longer(
        cols = starts_with("Steps"),
        names_to = "ActivityMinute",
        values_to = "Steps"
    ) %>%
    mutate(
        ActivityMinute = ActivityHour + minutes(as.integer(substr(ActivityMinute, 6, 7))), 
        ActivityHour = NULL,
        .before = "Steps"
    )

m2_daily_sleep <- read_csv(
    "datasets/fitbit_mturk_4.12.16-5.12.16/sleepDay_merged.csv",
    col_types = cols(
        SleepDay = col_datetime(format = "%m/%d/%Y %I:%M:%S %p")
    ),
)

#### TESTS
####
all_tests_passed <- TRUE

all_tibbles <- list(
    m1_daily_activity,
    m2_daily_activity,
    m1_heartrate_seconds,
    m2_heartrate_seconds,
    m1_hourly_calories,
    m2_hourly_calories,
    m1_minute_calories_n,
    m2_minute_calories_n,
    m1_hourly_intensities,
    m2_hourly_intensities,
    m1_minute_intensities_n,
    m2_minute_intensities_n,
    m1_hourly_steps,
    m2_hourly_steps,
    m1_minute_steps_n,
    m2_minute_steps_n,
    m1_minute_mets_n,
    m2_minute_mets_n,
    m1_minute_sleep, # has duplicates (525)
    m2_minute_sleep, # has duplicates (543)
    m1_weight_log,
    m2_weight_log,
    m2_daily_calories,
    m2_daily_intensities,
    m2_daily_steps,
    m2_minute_calories_n_v2,
    m2_minute_intensities_n_v2,
    m2_minute_steps_n_v2,
    m2_daily_sleep # has duplicates (29)
)

## Uncomment to check for duplication
# for (i in seq_along(all_tibbles)) {
#     tib_obj <- all_tibbles[[i]]
#     original_n_rows <- nrow(tib_obj)
#     distinct_n_rows <- nrow(distinct(tib_obj))
#     if (original_n_rows != distinct_n_rows) {
#         print(str_glue("Tibble at index {i} has {abs(original_n_rows - distinct_n_rows)} duplicate rows ({original_n_rows} v. {distinct_n_rows}):")) 
#     }
# }

## Uncomment to check if shared datasets of two months are structurally the same
# for (i in seq_along(shared_tibbles)) {
#     tibble_pair <- shared_tibbles[[i]]
#     passed <- compare_tibbles(tibble_pair[[1]], tibble_pair[[2]], i)
#     if (!passed) all_tests_passed <- FALSE
# }

## Uncomment to see if there are any NA values
# for (i in seq_along(all_tibbles)) {
#     map(all_tibbles[i], function(tib_obj) {
#         result <- sapply(tib_obj, function(x) sum(is.na(x)))
#         print(result)
#     })
# }

if (all_tests_passed) {
    print("All tests passed!")
} else {
    print("Some tests failed.")
}

#### TRANSFORMATIONS
####
# Remove duplication
m1_minute_sleep <- distinct(m1_minute_sleep)
m2_minute_sleep <- distinct(m2_minute_sleep)
m2_daily_sleep <- distinct(m2_daily_sleep)

# m2_minute_calories_n and m2_minute_calories_n_v2
# m2_minute_intensities_n and m2_minute_intensities_n_v2
# m2_minute_steps_n and m2_minute_steps_n_v2
# are each missing data from the other

m2_minute_calories_n <- union(m2_minute_calories_n, m2_minute_calories_n_v2)
m2_minute_intensities_n <- union(m2_minute_intensities_n, m2_minute_intensities_n_v2) 
m2_minute_steps_n <- union(m2_minute_steps_n, m2_minute_steps_n_v2)

# group m2_minute_calories_union and compare against m2_daily_calories
# group m2_minute_intensities_union and compare against m2_daily_intensities
# group m2_minute_steps_union and compare against m2_daily_steps

#### Build *_daily_activity / fill in the gaps

## Need for each month:
# daily calories
# daily intensities
# daily mets
# daily steps

## Other
# daily sleep

## Month 1 datasets (visual reminder)
#  - dailyActivity
#  - hourlyCalories
#  - hourlyIntensities
#  - hourlySteps
#  - minuteCaloriesNarrow
#  - minuteIntensitiesNarrow
#  - minuteMETsNarrow
#  - minuteSleep
#  - minuteStepsNarrow
#  - heartrate_seconds
#  - weightLogInfo

## Month 1 needs:
# daily calories - done
# daily mets - done
# daily steps - done
# daily intensities - done
# daily sleep - done

## Month 1 Calories
m1_daily_calories <- m1_minute_calories_n %>%
    mutate(
        ActivityDay = date(ActivityMinute),
        ActivityMinute = NULL
    ) %>%
    group_by(Id, ActivityDay) %>%
    summarise(Calories = sum(Calories)) %>%
    rename(ActivityDate = "ActivityDay")


## Month 1 METs // Divide by 10
m1_daily_mets <- m1_minute_mets_n %>%
    mutate(
        ActivityDay = date(ActivityMinute),
        ActivityMinute = NULL,
        METs = (METs / 10)
    ) %>%
    group_by(Id, ActivityDay) %>%
    summarise(METs = sum(METs >= 2.4)) %>%
    rename(ActivityDate = "ActivityDay")

## Month 1 Sleep
m1_daily_sleep <- m1_minute_sleep %>%
    rename(
        LogId = "logId",
        SleepDay = "date"
    ) %>%
    mutate(
        SleepDay = date(SleepDay)
    ) %>%
    group_by(Id, SleepDay) %>%
    summarize(
        TotalSleepRecords = n_distinct(LogId),
        TotalMinutesAsleep = sum(value == 1),
        TotalMinutesRestless = sum(value == 2),
        TotalTimeInBed = n(),
        Efficiency = (100 * TotalMinutesAsleep / TotalTimeInBed),
        Inefficiency = (100 * TotalMinutesRestless / TotalTimeInBed)
    ) %>%
    rename(ActivityDate = "SleepDay")

## Month 1 Steps
m1_daily_steps <- m1_minute_steps_n %>%
    mutate(
        ActivityDay = date(ActivityMinute),
        ActivityMinute = NULL
    ) %>%
    group_by(Id, ActivityDay) %>%
    summarise(StepTotal = sum(Steps)) %>%
    rename(ActivityDate = "ActivityDay")

## Month 1 Intensities
# 0 = Sedentary, 1 = Light, 2 = Moderate, 3 = Very Active
m1_daily_intensities <- m1_minute_intensities_n %>%
    mutate(
        ActivityDay = date(ActivityMinute),
        ActivityMinute = NULL
    ) %>%
    group_by(Id, ActivityDay) %>%
    summarise(
        SedentaryMinutes = sum(Intensity == 0),
        LightlyActiveMinutes = sum(Intensity == 1),
        FairlyActiveMinutes = sum(Intensity == 2),
        VeryActiveMinutes = sum(Intensity == 3)
    ) %>%
    rename(ActivityDate = "ActivityDay")

## Month 1 Data Collage
m1_daily_activity_final <- full_join(
    m1_daily_calories, m1_daily_mets,
    by = join_by(Id, ActivityDate)
) %>%
    # full_join(m1_daily_sleep, by = join_by(Id, ActivityDate)) %>% Omit sleep
    full_join(m1_daily_steps, by = join_by(Id, ActivityDate)) %>%
    full_join(m1_daily_intensities, by = join_by(Id, ActivityDate)) %>%
    rename(TotalSteps = "StepTotal")

## Month 2 datasets (visual reminder)
#  - dailyActivity
#  - dailyCalories
#  - dailyIntensities
#  - dailySteps
#  - sleepDay
#  - hourlyCalories
#  - hourlyIntensities
#  - hourlySteps
#  - minuteCaloriesNarrow
#  - minuteIntensitiesNarrow
#  - minuteMETsNarrow
#  - minuteSleep
#  - minuteStepsNarrow
#  - minuteCaloriesWide
#  - minuteIntensitiesWide
#  - minuteStepsWide
#  - heartrate_seconds
#  - weightLogInfo

# Month 2 needs:
# calories, intensities, and steps rebuilt to conform to unionized data
# daily mets
# daily sleep (needs sleep efficiency column)

## Month 2 Calories
m2_daily_calories <- m2_minute_calories_n %>%
    mutate(
        ActivityDay = date(ActivityMinute),
        ActivityMinute = NULL
    ) %>%
    group_by(Id, ActivityDay) %>%
    summarise(Calories = sum(Calories)) %>%
    rename(ActivityDate = "ActivityDay")

# Month 2 METs // Divide by 10
m2_daily_mets <- m2_minute_mets_n %>%
    mutate(
        ActivityDay = date(ActivityMinute),
        ActivityMinute = NULL,
        METs = (METs / 10)
    ) %>%
    group_by(Id, ActivityDay) %>%
    summarise(METs = sum(METs >= 2.4)) %>%
    rename(ActivityDate = "ActivityDay")

## Mother 2 Steps
m2_daily_steps <- m2_minute_steps_n %>%
    mutate(
        ActivityDay = date(ActivityMinute),
        ActivityMinute = NULL
    ) %>%
    group_by(Id, ActivityDay) %>%
    summarise(StepTotal = sum(Steps)) %>%
    rename(ActivityDate = "ActivityDay")

## Month 2 Intensities
m2_daily_intensities <- m2_minute_intensities_n %>%
    mutate(
        ActivityDay = date(ActivityMinute),
        ActivityMinute = NULL
    ) %>%
    group_by(Id, ActivityDay) %>%
    summarise(
        SedentaryMinutes = sum(Intensity == 0),
        LightlyActiveMinutes = sum(Intensity == 1),
        FairlyActiveMinutes = sum(Intensity == 2),
        VeryActiveMinutes = sum(Intensity == 3)
    ) %>%
    rename(ActivityDate = "ActivityDay")

## Month 2 Sleep
# value column 1 = asleep 2 = restless 3 = awake
# group by logId, use restless to calculate sleep efficiency
# m2_daily_sleep <- m2_daily_sleep %>%
#     mutate(
#         Efficiency = (100 * TotalMinutesAsleep / TotalTimeInBed)
#     ) %>%
#     rename(ActivityDate = "SleepDay")
m2_daily_sleep <- m2_minute_sleep %>%
    rename(
        LogId = "logId",
        SleepDay = "date"
    ) %>%
    mutate(
        SleepDay = date(SleepDay)
    ) %>%
    group_by(Id, SleepDay) %>%
    summarize(
        TotalSleepRecords = n_distinct(LogId),
        TotalMinutesAsleep = sum(value == 1),
        TotalMinutesRestless = sum(value == 2),
        TotalTimeInBed = n(),
        Efficiency = (100 * TotalMinutesAsleep / TotalTimeInBed),
        Inefficiency = (100 * TotalMinutesRestless / TotalTimeInBed)
    ) %>%
    rename(ActivityDate = "SleepDay")

## Month 2 Data Collage
m2_daily_activity_final <- full_join(
    m2_daily_calories, m2_daily_mets,
    by = join_by(Id, ActivityDate)
) %>%
    # full_join(m2_daily_sleep, by = join_by(Id, ActivityDate)) %>% Omit sleep
    full_join(m2_daily_steps, by = join_by(Id, ActivityDate)) %>%
    full_join(m2_daily_intensities, by = join_by(Id, ActivityDate)) %>%
    rename(TotalSteps = "StepTotal")

## Uncomment for a difference comparison between original and new activity sheets
# m1_daily_activity_trimmed <- m1_daily_activity %>%
#     mutate(
#         TotalDistance = NULL,
#         TrackerDistance = NULL,
#         LoggedActivitiesDistance = NULL,
#         VeryActiveDistance = NULL,
#         ModeratelyActiveDistance = NULL,
#         LightActiveDistance = NULL,
#         SedentaryActiveDistance = NULL,
#     )

# m2_daily_activity_trimmed <- m2_daily_activity %>%
#     mutate(
#         TotalDistance = NULL,
#         TrackerDistance = NULL,
#         LoggedActivitiesDistance = NULL,
#         VeryActiveDistance = NULL,
#         ModeratelyActiveDistance = NULL,
#         LightActiveDistance = NULL,
#         SedentaryActiveDistance = NULL,
#     )

# m1_daily_activity_final_trimmed <- m1_daily_activity_final %>% #nolint
#     mutate(
#         METs = NULL
#     )

# m2_daily_activity_final_trimmed <- m2_daily_activity_final %>% #nolint
#     mutate(
#         METs = NULL
#     )

# m1_diffs <- symdiff(m1_daily_activity_trimmed, m1_daily_activity_final_trimmed)
# m2_diffs <- symdiff(m2_daily_activity_trimmed, m2_daily_activity_final_trimmed)
# m1_daily_activity_unified <- union(m1_daily_activity_trimmed, m1_daily_activity_final_trimmed)
# view(m1_daily_activity_unified)
## A union of these datasets produces duplicates as
## a result of very small degree in calculation
## (For example, 1817 calories vs 1817.0229 calories)

all_daily_activity <- union(m1_daily_activity_final, m2_daily_activity_final) %>% 
    distinct(Id, ActivityDate, .keep_all = TRUE)

full_period_summary <- all_daily_activity %>%
    group_by(Id) %>%
    summarise(
        TotalDaysMonitored = n(),
        FirstReportingDate = min(ActivityDate),
        LastReportingDate = max(ActivityDate),
        TotalDaysInPeriod = abs(as.numeric(difftime(FirstReportingDate, LastReportingDate))), 
        PercentageDaysReporting = (100 * TotalDaysMonitored) / TotalDaysInPeriod, 
        #
        TotalCaloriesBurned = sum(Calories),
        AverageCaloriesBurned = mean(Calories),
        HighestCaloriesBurned = max(Calories),
        #
        TotalMETs = sum(METs, na.rm = TRUE),
        AverageMETs = mean(METs, na.rm = TRUE),
        HighestMETs = max(METs, na.rm = TRUE),
        #
        TotalSteps = sum(TotalSteps),
        AverageSteps = mean(TotalSteps),
        HighestSteps = max(TotalSteps),
        #
        TotalSedentaryMinutes = sum(SedentaryMinutes),
        AverageSedentaryMinutes = mean(SedentaryMinutes),
        HighestSedentaryMinutes = max(SedentaryMinutes),
        #
        TotalLightlyActiveMinutes = sum(LightlyActiveMinutes),
        AverageLightlyActiveMinutes = mean(LightlyActiveMinutes),
        HighestLightlyActiveMinutes = max(LightlyActiveMinutes),
        #
        TotalFairlyActiveMinutes = sum(FairlyActiveMinutes),
        AverageFairlyActiveMinutes = mean(FairlyActiveMinutes),
        HighestFairlyActiveMinutes = max(FairlyActiveMinutes),
        FairlyActiveBurnoutRatio = round(100 * (HighestFairlyActiveMinutes / TotalFairlyActiveMinutes), digits = 1), 
        #
        TotalVeryActiveMinutes = sum(VeryActiveMinutes),
        AverageVeryActiveMinutes = mean(VeryActiveMinutes),
        HighestVeryActiveMinutes = max(VeryActiveMinutes),
        VeryActiveBurnoutRatio = round(100 * (HighestVeryActiveMinutes / TotalVeryActiveMinutes), digits = 1), 
    ) %>%
    filter(Id != 2891001357) %>%
    # Some PercentageDaysReporting entries are > 100%,
    # likely due to a quirk in how days in the period are calculated
    mutate(
        PercentageDaysReporting = case_when(
            PercentageDaysReporting > 100 ~ 100,
            PercentageDaysReporting <= 100 ~ round(PercentageDaysReporting, digits = 1) 
        ),
        ReportingGroup = case_when(
            PercentageDaysReporting == 100 ~ "=100%",
            PercentageDaysReporting >= 90 & PercentageDaysReporting < 100 ~ "90-99%", 
            PercentageDaysReporting >= 80 & PercentageDaysReporting < 90 ~ "80-89%", 
            PercentageDaysReporting < 80 ~ "<80%"
        )
    )

participant_summary <- full_period_summary %>%
    summarise(
        NumberOfParticipants = n(),
        ShortestPeriodParticipation = min(TotalDaysInPeriod),
        LongestPeriodParticipation = max(TotalDaysInPeriod),
        PercantageFullyReporting = 100 * sum(PercentageDaysReporting == 100) / NumberOfParticipants 
    )
participant_pie <- full_period_summary %>%
    group_by(ReportingGroup) %>%
    summarise(
        fraction = n() / nrow(.),
    ) %>%
    mutate(
        ymax = cumsum(fraction),
        ymin = c(0, head(ymax, n = -1)),
        labelPosition = (ymax + ymin) / 2,
        label = paste0("(", round(100 * fraction, digits = 2), "%)"), 
    ) %>%
    ggplot(aes(
        ymax = ymax,
        ymin = ymin,
        xmax = 4,
        xmin = 3,
        fill = ReportingGroup
    )) +
    geom_rect() +
    geom_text(
        x = 2.2,
        aes(y = labelPosition, label = label, color = ReportingGroup, fontface = "bold"), 
        show.legend = FALSE,
        size = 8
    ) +
    scale_fill_brewer(palette = 3) +
    scale_color_brewer(palette = 3) +
    coord_polar(theta = "y") +
    xlim(c(-1, 4)) +
    theme_void() +
    theme(
        legend.title = element_text(size = 26),
        legend.key.size = unit(2, "cm"),
        legend.text = element_text(size = 22)
    ) +
    labs(fill = "Days Reporting")

all_weight_logs <- union(m1_weight_log, m2_weight_log)

first_measurement_dates <- all_weight_logs %>%
    select(Id, Date, WeightKg, BMI) %>%
    group_by(Id) %>%
    filter(
        Date == min(Date)
    ) %>%
    rename(
        FirstWeightKg = "WeightKg",
        FirstBMI = "BMI"
    )

last_measurement_dates <- all_weight_logs %>%
    select(Id, Date, WeightKg, BMI) %>%
    group_by(Id) %>%
    filter(
        Date == max(Date)
    ) %>%
    rename(
        LastWeightKg = "WeightKg",
        LastBMI = "BMI"
    )

all_measurement_changes <- first_measurement_dates %>%
    full_join(last_measurement_dates, by = join_by(Id)) %>%
    mutate(
        WeightKgChange = LastWeightKg - FirstWeightKg,
        BMIChange = LastBMI - FirstBMI
    )

all_weight_summary <- all_weight_logs %>%
    select(Id, Date, WeightKg, BMI) %>%
    group_by(Id) %>%
    summarise(
        TotalMeasurements = n()
    ) %>%
    full_join(all_measurement_changes, by = join_by(Id))

all_weight_changes <- all_weight_summary %>%
    filter(WeightKgChange >= .5 | WeightKgChange <= -.5)

weight_changes_graph <- all_weight_changes %>%
    ggplot(
        mapping = aes(
            x = as.character(Id),
            y = WeightKgChange,
            fill = as.factor(Id)
        ),
    ) +
    geom_bar(
        stat = "identity",
        width = .8,
    ) +
    labs(
        title = "Changes in Participant Weight (kg)",
        subtitle = "all changes > .5 kg",
        x = "Participant ID",
        y = "Weight Change (kg)"
    ) +
    geom_hline(yintercept = 0, lwd = 0.8, colour = "#272525") +
    theme(
        legend.position = "none",
        text = element_text(size = 24),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text.x.bottom = element_text(angle = 25, size = 20, face = "bold")
    )

overweight_participants_logs <- all_weight_summary %>%
    filter(FirstBMI > 25 & TotalMeasurements > 1) %>%
    select(Id) %>%
    left_join(all_weight_logs, by = join_by(Id))

bmi_v_measurements_graph <- overweight_participants_logs %>%
    ggplot(
        mapping = aes(
            x = Date,
            y = BMI,
            fill = "white",
            color = as.factor(Id),
            size = 3
        )
    ) +
    labs(
        title = "BMI Over Time (Overweight)",
        subtitle = "each point a BMI log"
    ) +
    geom_point() +
    theme(
        legend.position = "none",
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold")
    )

normal_weight_participants_logs <- all_weight_summary %>%
    filter(FirstBMI <= 25 & TotalMeasurements > 1) %>%
    select(Id) %>%
    left_join(all_weight_logs, by = join_by(Id))

### A graph of "healthy" BMI participants
normal_bmi_v_t_graph <- normal_weight_participants_logs %>%
    ggplot(
        mapping = aes(
            x = Date,
            y = BMI,
            fill = "white",
            color = as.factor(Id),
            size = 4
        )
    ) +
    labs(
        title = "BMI Over Time for BMI <= 25",
        subtitle = "each point a BMI log"
    ) +
    geom_point() +
    theme(
        legend.position = "none",
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold")
    )
# calories per step during light activity,
all_minute_calories <- union(m1_minute_calories_n, m2_minute_calories_n)
all_minute_steps <- union(m1_minute_steps_n, m2_minute_steps_n)

all_minute_light_activity <- union(m1_minute_intensities_n, m2_minute_intensities_n) %>% 
    filter(Intensity == 1)

all_minute_fairly_activity <- union(m1_minute_intensities_n, m2_minute_intensities_n) %>% 
    filter(Intensity == 2)

all_minute_very_activity <- union(m1_minute_intensities_n, m2_minute_intensities_n) %>% 
    filter(Intensity == 3)

all_minute_activity <- all_minute_light_activity %>%
    union(all_minute_fairly_activity) %>%
    union(all_minute_very_activity)

cal_v_step_v_light_activty <- all_minute_calories %>%
    full_join(all_minute_steps, by = join_by(Id, ActivityMinute)) %>%
    full_join(all_minute_light_activity, by = join_by(Id, ActivityMinute)) %>%
    drop_na() %>%
    group_by(Id) %>%
    summarise(
        TotalMinutes = n(),
        TotalSteps = sum(Steps),
        TotalCalories = sum(Calories)
    ) %>%
    filter(Id != 289100135)

cal_v_step_v_fairly_activty <- all_minute_calories %>%
    full_join(all_minute_steps, by = join_by(Id, ActivityMinute)) %>%
    full_join(all_minute_fairly_activity, by = join_by(Id, ActivityMinute)) %>%
    drop_na() %>%
    group_by(Id) %>%
    summarise(
        TotalMinutes = n(),
        TotalSteps = sum(Steps),
        TotalCalories = sum(Calories)
    ) %>%
    filter(Id != 289100135)

cal_v_step_v_very_activty <- all_minute_calories %>%
    full_join(all_minute_steps, by = join_by(Id, ActivityMinute)) %>%
    full_join(all_minute_very_activity, by = join_by(Id, ActivityMinute)) %>%
    drop_na() %>%
    group_by(Id) %>%
    summarise(
        TotalMinutes = n(),
        TotalSteps = sum(Steps),
        TotalCalories = sum(Calories)
    ) %>%
    filter(Id != 289100135)

cal_v_step_v_all_activty <- all_minute_calories %>%
    full_join(all_minute_steps, by = join_by(Id, ActivityMinute)) %>%
    full_join(all_minute_activity, by = join_by(Id, ActivityMinute)) %>%
    drop_na() %>%
    group_by(Id) %>%
    summarise(
        TotalMinutes = n(),
        TotalSteps = sum(Steps),
        TotalCalories = sum(Calories)
    ) %>%
    filter(Id != 289100135)

## Join with BMI
all_weighted_participants <- all_weight_summary %>%
    select(Id, LastBMI) %>%
    rename(BMI = "LastBMI")

light_activity_v_bmi <- cal_v_step_v_light_activty %>%
    left_join(all_weighted_participants, by = join_by(Id)) %>%
    drop_na()

fairly_activity_v_bmi <- cal_v_step_v_fairly_activty %>%
    left_join(all_weighted_participants, by = join_by(Id)) %>%
    drop_na()

very_activity_v_bmi <- cal_v_step_v_very_activty %>%
    left_join(all_weighted_participants, by = join_by(Id)) %>%
    drop_na()

all_activity_v_bmi <- cal_v_step_v_all_activty %>%
    left_join(all_weighted_participants, by = join_by(Id)) %>%
    drop_na()

light_v_bmi_graph <- light_activity_v_bmi %>%
    ggplot(
        aes(
            x = BMI,
            y = TotalSteps,
            fill = as.factor(Id)
        )
    ) +
    geom_bar(
        stat = "identity",
        width = .8
    ) +
    # scale_fill_brewer(palette = 3) +
    # scale_color_brewer(palette = 3) +
    labs(
        title = "Light Activity Vs. BMI",
        y = "Total Light Activity Steps"
    ) +
    theme(
        legend.position = "none",
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold")
    )

fairly_v_bmi_graph <- fairly_activity_v_bmi %>%
    ggplot(
        aes(
            x = BMI,
            y = TotalSteps,
            fill = as.factor(Id)
        )
    ) +
    geom_bar(
        stat = "identity",
        width = .8
    ) +
    # scale_fill_brewer(palette = 3) +
    # scale_color_brewer(palette = 3) +
    labs(
        title = "Fairly Active Steps Vs. BMI",
        y = "Total Fairly Active Steps"
    ) +
    theme(
        legend.position = "none",
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold")
    )

very_v_bmi_graph <- very_activity_v_bmi %>%
    ggplot(
        aes(
            x = BMI,
            y = TotalSteps,
            fill = as.factor(Id)
        )
    ) +
    geom_bar(
        stat = "identity",
        width = .8
    ) +
    # scale_fill_brewer(palette = 3) +
    # scale_color_brewer(palette = 3) +
    labs(
        title = "Very Active Steps Vs. BMI",
        y = "Total Very Active Steps"
    ) +
    theme(
        legend.position = "none",
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold")
    )

light_steps_v_min_v_bmi_graph <- light_activity_v_bmi %>%
    ggplot(
        aes(
            x = BMI,
            y = (TotalSteps / TotalMinutes),
            fill = as.factor(Id)
        )
    ) +
    geom_bar(
        stat = "identity",
        width = .8
    ) +
    # scale_fill_brewer(palette = 3) +
    # scale_color_brewer(palette = 3) +
    labs(
        title = "Light Steps / Minute Vs. BMI",
        y = "Steps/minute"
    ) +
    theme(
        legend.position = "none",
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold")
    )

active_steps_v_min_v_bmi_graph <- very_activity_v_bmi %>%
    ggplot(
        aes(
            x = BMI,
            y = (TotalSteps / TotalMinutes),
            fill = as.factor(Id)
        )
    ) +
    geom_bar(
        stat = "identity",
        width = .8
    ) +
    # scale_fill_brewer(palette = 3) +
    # scale_color_brewer(palette = 3) +
    labs(
        title = "Very Active Steps / Minute Vs. BMI",
        y = "Steps/minute"
    ) +
    theme(
        legend.position = "none",
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold")
    )

## People with lower BMIs tend to exercise significantly more
fairly_steps_v_min_v_bmi_graph <- fairly_activity_v_bmi %>%
    ggplot(
        aes(
            x = BMI,
            y = (TotalSteps / TotalMinutes),
            fill = as.factor(Id)
        )
    ) +
    geom_bar(
        stat = "identity",
        width = .8
    ) +
    # scale_fill_brewer(palette = 3) +
    # scale_color_brewer(palette = 3) +
    labs(
        title = "Fairly Active Steps / Minute Vs. BMI",
        y = "Steps/minute"
    ) +
    theme(
        legend.position = "none",
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold")
    )

all_daily_mets <- union(m1_daily_mets, m2_daily_mets) %>%
    filter(Id != 2891001357 & METs > 0)

peak_mets_dates <- all_daily_mets %>%
    filter(METs == max(METs) & Id != 2891001357)

all_daily_mets <- all_daily_mets %>%
    inner_join(peak_mets_dates, by = join_by(Id))

prepeak_periods <- all_daily_mets %>%
    filter(ActivityDate.x <= ActivityDate.y) %>%
    select(-ActivityDate.y, -METs.y) %>%
    rename(METs = "METs.x", ActivityDate = "ActivityDate.x")

postpeak_periods <- all_daily_mets %>%
    filter(ActivityDate.x > ActivityDate.y) %>%
    select(-ActivityDate.y, -METs.y) %>%
    rename(METs = "METs.x", ActivityDate = "ActivityDate.x")

prepeak_summary <- prepeak_periods %>%
    group_by(Id) %>%
    summarise(
        PrepeakPeriodDays = n(),
        PrepeakTotalMETs = sum(METs),
        PrepeakAverage = PrepeakTotalMETs / PrepeakPeriodDays
    )

postpeak_summary <- postpeak_periods %>%
    group_by(Id) %>%
    summarise(
        PostpeakPeriodDays = n(),
        PostpeakTotalMets = sum(METs),
        PostpeakAverage = PostpeakTotalMets / PostpeakPeriodDays
    )

burnout_summary <- prepeak_summary %>%
    full_join(postpeak_summary, by = join_by(Id)) %>%
    filter(PostpeakPeriodDays >= 7) %>%
    filter(PrepeakAverage > (PostpeakAverage * 1.33)) # 33% Decline factor

burnout_daily_mets <- all_daily_mets %>%
    inner_join(burnout_summary, by = join_by(Id)) %>%
    rename(METs = "METs.x", ActivityDate = "ActivityDate.x")

burnout_daily_mets$week_num <- strftime(burnout_daily_mets$ActivityDate, format = "%V") 

## Use this...
burnout_summary_temp <- burnout_summary %>%
    select(Id)

burnout_weekly_mets_temp <- burnout_daily_mets %>%
    select(Id, ActivityDate, METs, week_num) %>%
    group_by(Id, week_num) %>%
    summarise(
        ActivityDate = min(ActivityDate),
        AverageMETs = mean(METs)
    ) %>%
    mutate(
        week_num = NULL,
        doColor = FALSE
    )

peak_mets_dates_temp <- peak_mets_dates %>%
    rename(AverageMETs = "METs") %>%
    right_join(burnout_summary_temp, by = join_by(Id)) %>%
    mutate(
        doColor = TRUE
    )

burnout_weekly_graph_data <- union(burnout_weekly_mets_temp, peak_mets_dates_temp) 

burnout_graph <- burnout_weekly_graph_data %>%
    ggplot(
        aes(
            x = ActivityDate,
            y = AverageMETs,
            color = as.factor(Id),
            fill = "Burnout Risks"
        )
    ) +
    geom_line(linewidth = 2) +
    geom_point(shape = 21, fill = "white", size = 5, stroke = 2) + # Base
    geom_point(aes(color = as.factor(Id))) + # Dotted center
    geom_point(
        data = peak_mets_dates_temp,
        aes(
            x = ActivityDate,
            y = AverageMETs,
            color = as.factor(Id),
            size = 8,
            stroke = 1,
            alpha = 1
        )
    ) +
    labs(
        title = "Weekly Average METs",
        subtitle = "as a predictor of burnout",
        fill = "Burnout Risks",
        y = "METs",
        x = ""
    ) +
    theme(
        legend.position = "none",
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.y.left = element_text(size = 20, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold"),
        legend.title = element_text(size = 24, face = "bold"),
        legend.text = element_text(size = 16, face = "bold"),
    )

general_peaks_graph <- peak_mets_dates %>%
    ggplot(
        aes(
            x = ActivityDate,
            y = METs,
            color = as.factor(Id),
            fill = "Burnout Risks"
        )
    ) +
    geom_point(shape = 21, fill = "white", size = 5, stroke = 2) + # Base
    geom_point(aes(color = as.factor(Id))) + # Dotted center
    labs(
        title = "Time of Peak METs",
        subtitle = "each point a participant",
        y = "METs",
        x = ""
    ) +
    scale_fill_brewer(palette = 3) +
    # scale_color_brewer(palette = 3) +
    theme(
        legend.position = "none",
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.y.left = element_text(size = 20, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold"),
        legend.title = element_text(size = 24, face = "bold"),
        legend.text = element_text(size = 16, face = "bold"),
    )

burnout_peaks_graph <- peak_mets_dates_temp %>%
    ggplot(
        aes(
            x = ActivityDate,
            y = AverageMETs,
            color = as.factor(Id),
            fill = "Burnout Risks"
        )
    ) +
    geom_point(shape = 21, fill = "white", size = 5, stroke = 2) + # Base
    geom_point(aes(color = as.factor(Id))) + # Dotted center
    labs(
        title = "Time of Peak METs",
        subtitle = "each point a participant at risk of burnout",
        y = "METs",
        x = ""
    ) +
    scale_fill_brewer(palette = 3) +
    theme(
        legend.position = "none",
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.y.left = element_text(size = 20, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold"),
        legend.title = element_text(size = 24, face = "bold"),
        legend.text = element_text(size = 16, face = "bold"),
    )

all_heartrate_seconds <- union(m1_heartrate_seconds, m2_heartrate_seconds) %>%
    rename(ActivityMinute = "Time", Heartrate = "Value")

## Reminder:
# all_minute_light_activity
# all_minute_fairly_activity
# all_minute_very_activity
# all_minute_activity

all_minute_sedentary_activity <- union(m1_minute_intensities_n, m2_minute_intensities_n) %>% 
    filter(Intensity == 0) %>%
    mutate(Intensity = NULL)

all_minute_sleep <- union(m1_minute_sleep, m2_minute_sleep) %>%
    select(Id, date) %>%
    rename(ActivityMinute = "date") %>%
    mutate(ActivityMinute = ActivityMinute + seconds(30))

all_sedentary_no_sleep <- setdiff(all_minute_sedentary_activity, all_minute_sleep) 

## To prevent definition clobbering
all_daily_mets_temp <- union(m1_daily_mets, m2_daily_mets) %>%
    filter(Id != 2891001357 & METs > 0)

heartrate_collage_sedentary <- all_heartrate_seconds %>%
    inner_join(
        all_sedentary_no_sleep,
        by = join_by(Id, ActivityMinute)
    ) %>%
    mutate(
        ActivityDate = date(ActivityMinute),
        ActivityMinute = NULL
    ) %>%
    group_by(Id, ActivityDate) %>%
    summarise(
        AverageHeartrate = round(mean(Heartrate))
    ) %>%
    inner_join(
        all_daily_mets_temp,
        by = join_by(Id, ActivityDate)
    )

heartrate_collage_period <- heartrate_collage_sedentary

heartrate_collage_period$month_num <- strftime(heartrate_collage_period$ActivityDate, format = "%m") 

heartrate_collage_period <- heartrate_collage_period %>%
    group_by(Id, month_num) %>%
    summarise(
        AverageHeartrate = round(mean(AverageHeartrate)),
        AverageMETs = round(mean(METs))
    ) %>%
    mutate(
        month_num = NULL
    )
heartrate_graph <- heartrate_collage_period %>%
    ggplot(
        mapping = aes(
            x = AverageMETs,
            y = AverageHeartrate,
            fill = "white",
            color = as.factor(Id),
            size = 3
        )
    ) +
    labs(
        title = "Heartrate vs. Activity Levels",
        subtitle = "during sedentary, non-sleeping hours",
        y = "Average Daily Heartrate",
        x = "Average Daily METs"
    ) +
    geom_point() +
    theme(
        legend.position = "none",
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.y.left = element_text(size = 20, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold")
    )

m1_daily_sleep_temp <- m1_daily_sleep %>%
    select(
        Id,
        ActivityDate,
        TotalMinutesAsleep,
        TotalMinutesRestless,
        Efficiency,
        Inefficiency
    )

m2_daily_sleep_temp <- m2_daily_sleep %>%
    select(
        Id,
        ActivityDate,
        TotalMinutesAsleep,
        TotalMinutesRestless,
        Efficiency,
        Inefficiency
    )

all_daily_sleep <- union(m1_daily_sleep_temp, m2_daily_sleep_temp) %>%
    group_by(Id) %>%
    summarise(
        AverageMinutesAsleep = mean(TotalMinutesAsleep),
        AverageMinutesRestless = mean(TotalMinutesRestless),
        AverageEfficiency = mean(Efficiency),
        AverageInefficiency = mean(Inefficiency)
    )

heart_sleep_collage <- inner_join(
    all_daily_sleep, heartrate_collage_period,
    by = join_by(Id)
)

sleep_efficiency_v_activity <- heart_sleep_collage %>%
    ggplot(
        mapping = aes(
            x = AverageEfficiency,
            y = AverageMETs,
            fill = "white",
            color = as.factor(Id),
            size = 3
        )
    ) +
    labs(
        title = "Activity vs. Sleep Efficiency",
        subtitle = "Efficiency = (Minutes Asleep / Minutes in Bed)",
        y = "Activity Levels",
        x = "Efficiency (%)"
    ) +
    geom_point() +
    theme(
        legend.position = "none",
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.y.left = element_text(size = 20, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold")
    )

sleep_inefficiency_v_activity <- heart_sleep_collage %>%
    ggplot(
        mapping = aes(
            x = AverageInefficiency,
            y = AverageMETs,
            fill = "white",
            color = as.factor(Id),
            size = 3
        )
    ) +
    labs(
        title = "Activity vs. Sleep Efficiency",
        subtitle = "Infficiency = (Minutes Restless / Minutes in Bed)",
        y = "Activity Levels",
        x = "Inefficiency (%)"
    ) +
    geom_point() +
    theme(
        legend.position = "none",
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.y.left = element_text(size = 20, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold")
    )

time_asleep_v_activity <- heart_sleep_collage %>%
    ggplot(
        mapping = aes(
            x = AverageMETs,
            y = AverageMinutesAsleep,
            fill = "white",
            color = as.factor(Id),
            size = 3
        )
    ) +
    labs(
        title = "Activity vs. Time Asleep",
        subtitle = "total minutes asleep (not just in bed)",
        x = "Activity Levels",
        y = "Time Asleep",
    ) +
    geom_point() +
    theme(
        legend.position = "none",
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.y.left = element_text(size = 20, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold")
    )

m1_minute_mets_temp <- m1_minute_mets_n %>%
    mutate(METs = METs / 10) # why Fitabase. just why

m2_minute_mets_temp <- m2_minute_mets_n %>%
    mutate(METs = METs / 10)

all_active_minute_mets <- union(m1_minute_mets_temp, m2_minute_mets_temp)

hourly_mets_summary <- all_active_minute_mets %>%
    filter(METs >= 2.4) %>%
    mutate(
        Minutes = as_datetime((hour(ActivityMinute) * 60 * 60) + minute(ActivityMinute) * 60) 
    ) %>%
    group_by(Id, Minutes) %>%
    summarise(
        AverageMETs = mean(METs)
    )

# max_mets <- hourly_mets_summary %>%
#     summarise(
#         MaxAverageMETs = max(AverageMETs)
#     )

hourlys_mets_by_id_graph <- hourly_mets_summary %>%
    ggplot(aes(
        x = Minutes,
        y = AverageMETs
    )) +
    geom_bin_2d() +
        labs(
        fill = "Count",
        title = "Intensity of Activity Throughout Day",
        subtitle = "for light and greater intensities",
        x = "Time of Day",
        y = "METs"
    ) +
    scale_x_datetime(
        date_labels = "%H:%M",
    ) +
    theme(
        text = element_text(size = 20),
        plot.title = element_text(size = 32, face = "bold"),
        axis.title.x = element_text(size = 28, face = "bold"),
        axis.title.y = element_text(size = 28, face = "bold"),
        axis.text.y.left = element_text(size = 20, face = "bold"),
        axis.text.y.right = element_text(size = 20, face = "bold"),
        axis.text.x.bottom = element_text(size = 22, face = "bold"),
    ) +
    coord_polar(theta = "x", clip = "off") +
    ylim(c(2.5, 13))

```

## Analysis

The final collage shows that there are thirty-five unique identifiers (**Id**). It might be assumed that each one matches one participant, but since only thirty participants are said to be in the study, it is more likely that the Ids pair with a specific device and that some participants simply had more than one device. With no way of knowing which device belonged to who, this analysis will treat each device as if it were a participant.

One participant (ID 2891001357) unfortunately only participated for one day. Some of their data has been removed for this study.

### The Participants

85% of participants committed fully to the program, reporting every day! Plus, the lowest percentage of days reported was only 85%.

Note, it's hard to say if the high reporting rate is a sign of enthusiasm for the device. It could be the people enthusiastic about their own health are drawn to using such devices, or that the participants really wanted that $10 provided through Mechanical Turk.

A number of users also did not wear their Fitbit all day, noting some incomplete records. This may indicate a low tolerance for the device, or that the device may contain an allergen. Bellabeat may want to trial users for long-term comfort.

```{r}
participant_pie
```

Of all participants, only thirteen reported their weight, and of those reporting, only four saw a change greater than half a kilogram over the study period.

```{r}
# BMI
weight_changes_graph
```

At least nine participants are [likely to be overweight (BMI > 25) with two likely to be obese](https://www.cdc.gov/obesity/basics/adult-defining.html#:~:text=If%20your%20BMI%20is%20less,falls%20within%20the%20obesity%20range.). Assuming these participants had weight loss targets, we can compare their change in weight against the number of times they reported. Frequent weigh-ins [may correlate](https://www.webmd.com/fitness-exercise/what-to-know-about-weighing-yourself#:~:text=Daily%20weigh%2Dins.&text=Research%20shows%20that%20people%20who,weigh%20in%20once%20a%20week.&text=Another%20benefit%20is%20that%20you,small%20losses%20throughout%20the%20week) with weight loss success.

```{r}
bmi_v_measurements_graph
```

Participants who measured their weight showed a very slight decrease in weight by the end of the study. Most participants reported their weight under three times. One poor individual saw an increase in four kilograms in three days and then never reported again.

It is hard to separate weight loss from natural variations in weight. Significant weight loss is achieved over longer periods than the two months in the study. For example, one participant weighed more than once every other day. It is unknown if they were conscientious of their weight or if they had a weight a loss target. Nonetheless, their weight fluctuated over a narrow range.

```{r}
normal_bmi_v_t_graph
```

### Activiy Levels

There is a relationship between BMI and levels activity. For this study and for this analysis, the level of activity is measured in both steps and [Metabolic Equivalents (METs)](https://www.healthline.com/health/what-are-mets#definition). A MET is the ratio of how over active you are over the amount of energy you use while at rest. A MET of 1.0 is how much energy you burn while at seated and doing nothing. Meanwhile, walking, for example, might be between 2.0 and 4.0 METs, while bicycling at 10 mph is closer to 8.0. The METs counted for this analysis are those recorded to be at least 2.4, indicating the upperend of the Light Activity category.

(Technical note: This measurement is derived from the weight of the participant and oxygen consumption. More intense activities require more oxygen, up to 3.5ml o2/ (1kg/1min). Additionally, the MET values housed at [Fitabase, the warehouse point for this data,](https://www.fitabase.com/media/1930/fitabasedatadictionary102320.pdf) multiples MET values by 10, so MET values have been divided by ten for input in this analysis.)

The American Heart Association [recommends 150 minutes of moderate activity](https://www.heart.org/en/healthy-living/fitness/fitness-basics/aha-recs-for-physical-activity-in-adults) per week. Provided a 4 mph walk is moderate activity (MET ~4.0) to most, that is a fitness goal of between 450 and 600 MET minutes per week.

```{r}
light_v_bmi_graph
```

The data is extremely limited, but individuals with a higher BMI show a marked decline in physical activity.

```{r}
fairly_v_bmi_graph
```

And this holds true across all categories of physical activity. One participant, however, is quite the performer, with nearly half a million very active steps.

```{r}
very_v_bmi_graph
```

During light activity, participants show comparable speeds.

```{r}
light_steps_v_min_v_bmi_graph
```

Interestingly, the difference in the number of steps per minute narrows across BMIs during intense activity. This suggests that, rather than being lazy, heavier people complete gait cycles at the same rate, but do not enjoy the same acceleration.

```{r}
active_steps_v_min_v_bmi_graph
```

Individuals with lower BMIs tend to exercise significantly more, however. The relationship between obesity and a lack of physical activity is likely to be a compounding issue.

```{r}

fairly_steps_v_min_v_bmi_graph
```

Some participants may have experienced burnout. Crudely, participants at risk of burnout were identified by noting the day of peak physical activity then measuring levels of activity in the days before and after. If there was at least a 33% decline in physical activity over the following week, the participant may be in a stage of burnout. These participants may be engaged in more physical activity that they can tolerate (either physically or mentally) or can allow time for.

As a whole, participants reach the individual peak levels activity at different junctures in time.

```{r}
general_peaks_graph
```

However, participants at risk of burnout tend to peak early...

```{r}
burnout_peaks_graph
```

... followed by a general decline in activity.

```{r}
burnout_graph
```

### Heartrate

The study also monitored participant heartrate. Higher activity levels are associated with a lower heart rate--activity trains the heart, making it stronger, therefore [making each beat more effective at circulating blood](https://www.health.harvard.edu/heart-health/is-a-low-heart-rate-worrisome#:~:text=Exercise%20strengthens%20the%20heart%20muscle,fit%20person%20while%20at%20rest.). This analysis looks at participant heartrates during sedentary, wakeful activities.

```{r}
heartrate_graph
```

The data generally shows more active participants clustered in the 60-75 beats-per-minute range. It's likely that the effects of good health are compounding: a stronger heart means you are able to engage in more physically intense activities for a longer period of time, which in turn leads to an even stronger heart.

### Sleep

Higher activity levels are associated with better sleep. Sleep efficiency is the number of minutes you spend sleeping vs. the number of minutes you spend just lying in bed. Active participants tend to spend more time sleeping while in bed.

```{r}
sleep_efficiency_v_activity
```

Likewise, less active individuals tend to spend more time restless in bed (unproductive in sleep, such as tossing and turning, rather than merely being awake.)

```{r}
sleep_inefficiency_v_activity
```

More active individuals often sleep longer as well.

```{r}
time_asleep_v_activity
```

### Periods of Activity

A heatmap of incidents of particular activity intesities at different times throughout the day shows that most activity is light activity that occurs roughly between 0900 and 1900. More intense activity can occur at any other time of the day, with the most extreme outlier (~12.5 METs) occuring at midnight. Participants in the study may show a varied schedule for activity or live across different parts of the world.

```{r}
hourlys_mets_by_id_graph
```

## Recommendations

### Inclusivity
At the time of writing, many cultures are undergoing a shift in attitudes towards gender and sexuality, generally becoming more inclusive in their definition of a woman. Bellabeat could benefit by considering less strongly feminine presenting women and non-cisgender individuals in designing their features and marketing materials.

### New Health Metrics
Rather than merely record heartrate, the Bellabeat device could identify possible unhealthy or symptomatic heart rhythms and report them to the wearer. This data does not appear to be in the Fitbit metrics.

### Encourage Healthy Habits
Bellabeat may want to make the device considerate of its users health needs and abilities. The definition of fitness varies across age-groups and physical capabilities. The Bellabeat device can help identify the needs of its users and make suggestions for the kinds of activities most likely to improve their health. Additionally, it can offer words of encouragement and understanding for users who have underperformed and help them reframe their mindset to be forward-thinking.

### Identify Burnout
Some participants may push themselves too hard. Identifying these users early may allow Bellabeat the chance to remedy their relationship with physical activity before the decline becomes a fixture in their health.

### Game-ify Participant Health
Programs that have gaming elements are more likely to keep user engagement. Good health doesn't necessarily need to be a game, but scoring participants and giving them milestone awards may increase their enjoyment. Duolingo has achieved a lot of success through this model while still teaching its users a new language. *Steps*, *METs*, and *Distance* may be good metrics to center their scoring system around. Bellabeat may want to look to Nintendo's use of *Miis* for inspiration, which are user-created avatars that can interact with other Miis across different Nintendo products.

### Tools to Track Metrics
Participants may be interested in tracking their own metrics, with results from analysis in the way that Spotify does its year-end Spotified Wrapped program (although this would hopefully be more frequent than once per year.)

## Final Notes
A big thank you to Laimis Andrijauskas and SALAH on Kaggle for their community contributions surrounding these datasets!